Denotational Semantics of Gradual Typing using Synthetic Guarded Domain Theory

### 2024-01-26 paper-new/intro.tex
more intro

An alternative approach, which we investigate in this paper, is provided by
\emph{synthetic guarded domain theory}\cite{birkedal-mogelberg-schwinghammer-stovring2011}.
The techniques of synthetic guarded domain theory allow us to internalize the
step-indexed reasoning normally required in logical relations proofs of graduality,
ultimately allowing us to specify the logical relation in a manner that looks nearly
identical to a typical, non-step-indexed logical relation.

@@

The main contribution of this work is a categorical and denotational
semantics for gradually typed langauges that models not just the term
language but the graduality property as well.
\begin{enumerate}
\item First, we give a rational reconstruction \max{todo: more shit.}
\end{enumerate}

>>>>

\max{the stuff below is out of date}
An alternative approach, which we investigate in this paper, is provided by
\emph{synthetic guarded domain theory}\cite{birkedal-mogelberg-schwinghammer-stovring2011}.
The techniques of synthetic guarded domain theory allow us to internalize the
step-indexed reasoning normally required in logical relations proofs of graduality,
ultimately allowing us to specify the logical relation in a manner that looks nearly
identical to a typical, non-step-indexed logical relation.

@@

The main contribution of this work is a categorical and denotational
semantics for gradually typed langauges that models not just the term
language but the graduality property as well.
\begin{enumerate}
\item First, we give a simple abstract categorical model of GTT using CBPV double categories.
\item Next, we modify this semantics to develop reflexive graph- and
  double categorical models that abstract over the details of
  step-indexed models, and provide a method for constructing such models.
\item We instantiate the abstract construction to provide a concrete
  semantics in informal guarded type theory which supports the
  addition of arbitrary algebraic effects.
\item We prove that the resulting denotational model is
  \emph{adequate} for the graduality property: a closed term precision
  $M \ltdyn N : \nat$ has the expected semantics, that $M$ errors or
  $M$ and $N$ have the same extensional behavior.
\end{enumerate}
